---
title: "41-modeling-forecast"
output: html_notebook
---

The purpose of this notebook is to use time series modeling to predict the *number* of permits for each desired subtypes for Nashville. The subtypes can in principle be built of off any number of categorical columns. We have chosen in general to use `comm_v_res` and `project_type`, which take on values of "commercial"/"residential" and "construction"/"demolition", respectively.

```{r Load}
source(knitr::purl("30-feature-engineering.Rmd", output="tmp-41.R", quiet=TRUE))
fs::file_delete("tmp-41.R")
p_load(forecast)
```

# Main forecast function

```{r forecast permits by subtype}
forecast_npermits_sub <- function(df, forecast_cols, forecast_yrs = 6, df_start_yr = 2016, df_start_mo = 07, freq = 12, make_plot = FALSE) {
  #' Forecast npermits by subtype
  #'
  #' Create a time-series forecast of the number of permits that will "occur" in the next couple years. The intended use case is to forecast the number of permits that will be issued, though this depends on the user supplying a dataframe of issued-only permits. This can be done by using the function `load_forecast_df`. In principle, forecasts could be constructed for other cities, but at the moment, `load_forecast_df` will only load Nashville permits. The forecasts returned are Gaussian. The returned tibble has several columns of interest: 1. count - forecasted number of permits of specific subtype; 2. sd - standard deviation of Gaussian forecast; 3. half_iqr - half of the inner quartile range; 4. half_i90 - half of the inner 90%. These values are intended for sampling and plotting desired confidence intervals of the forecasts.
  #'
  #' @param df (tibble) Tibble containing historical permits (output of `load_forecast_df`).
  #' @param forecast_cols (chr array) Columns in `df` used to create the subtypes.
  #' @param forecast_yrs  (num, default=6) Number of years to forecast. Forecast starts at the end of the data in df.
  #' @param df_start_yr (num, default=2016) First calendar year in `df` to use when building forecast.
  #' @param df_start_mo (num, default=07) First month in `df` to use when building forecast.
  #' @param freq (num, default=12) Frequency of forecast. Either 12 (monthly) or 1 (yearly).
  #' @param make_plot (bool, default=FALSE) If TRUE, show plots of forecast.
  #'
  #' @return (tibble) Tibble containing both historical numbers of permits and future forecasted numbers
  #' @export
  #'
  #' @examples
  #' load_forecast_df() %>% 
  #'   forecast_npermits_sub(forecast_cols = c("comm_v_res","project_type"), freq = 12, make_plot = TRUE)
  
  # set up start list for forecast function
  start_list <- c(df_start_yr, df_start_mo)
  start_list <- start_list[!is.na(start_list)]
  
  # Create grouping categories based on frequency
  if(freq==12){
    group_cats <- c("yrmonth", forecast_cols)
  }
  else if(freq==1){
    group_cats <- c("fy", forecast_cols)
  }
  else{
    stop("Frequency must either be 12 (monthly) or 1 (yearly)")
  }
  
  # grouped df with counts and type category
  df_counts <- df %>% 
    group_by(across(all_of(group_cats))) %>% 
    summarise(count = n()) %>% 
    ungroup() %>% 
    unite(type, all_of(forecast_cols), sep=" ")
  
  # set up unique values of type; one forecast per type
  unique_types <- unique(df_counts$type)
  
  # map forecast function across types and merge resulting dataframe
  forecasts <- unique_types %>%
    map(., ~forecast_by_type(., df_counts, start_list = start_list, yrs=forecast_yrs, freq=freq, make_plot = make_plot)) %>%
    bind_rows() %>%
    arrange(date, type) %>%
    mutate(sd = count-q_16) %>%
    mutate(half_iqr = count - q_25) %>% 
    mutate(half_i90 = count - q_05) %>% 
    select(date, type, count, sd, half_iqr, half_i90) %>%
    mutate(data = "prediction")
  
  # add sd and data 
  df_counts <- df_counts %>% 
    mutate(sd = 0, half_iqr = 0, half_i90 = 0, data = "observation")
  
  # probably is a more succinct way to do this using lubridate, but oh well
  if(freq == 12){

    forecasts <- forecasts  %>%
      mutate(date = as.Date(paste("01",date,sep="-"), format = "%d-%b %Y")) %>%
      mutate(year = lubridate::year(date), month=lubridate::month(date,label=TRUE)) %>%
      select(year, month, type, count, sd, half_iqr, half_i90, data)

    df_counts <- df_counts %>%
      mutate(date = as.Date(str_c(yrmonth, "01"), "%Y%m%d")) %>%
      mutate(year = lubridate::year(date), month=lubridate::month(date, label=TRUE)) %>% 
      select(year, month, type, count, sd, half_iqr, half_i90, data)
  }
  if(freq == 1){
    df_counts <- df_counts  %>%
      mutate(year = as.numeric(fy)) %>%
      select(year, type, count, sd, half_iqr, half_i90, data)
    
    forecasts <- forecasts  %>%
      mutate(year = as.numeric(date)) %>% 
      select(year, type, count, sd, half_iqr, half_i90, data)
  }

  # combine dfs and separate columns
  results <- df_counts %>%
    bind_rows(forecasts) %>%
    separate(type, forecast_cols) %>% 
    as.tibble()
  
  return(results)
}
```

# Forecast specific subtype

```{r make forecast for a specific type}
forecast_by_type <- function(type_val, df, start_list, yrs, freq=12, make_plot=FALSE) {
  #' Forecast npermits for a given subtype
  #'
  #' Forecast the number of permits for a given subtype (e.g. "commercial construction"). This function is not intended to be run by itself, but rather is called within `forecast_npermits_sub`. 
  #'
  #' @param type_val (char) String specifying subtype (e.g. "commercial construction")
  #' @param df (dataframe) Dataframe containing time, subtype, and counts.
  #' @param start_list (num array) Year, month to start forecast. If yearly forecasts are desired, set month to NA.
  #' @param yrs (num) Number of years to forecast. Forecast starts at the end of the data in df.
  #' @param freq (num, default=12) Frequency of forecast. Either 12 (monthly) or 1 (yearly).
  #' @param make_plot (bool, default=FALSE) If TRUE, show plots of forecast.
  #'
  #' @return dataframe containing forecasts for this subtype for desired timeframe.
  #' @export
  #'
  #' @examples See source code of `forecast_npermits_sub`. Not intended as a standalone function.
  
  start_list <- start_list[!is.na(start_list)]
  
  count <- df %>% 
    filter(type==type_val) %>% 
    select(count)
  
  # number of forecasts
  nforecasts <- freq*yrs
  
  # forecast
  type_ts <- ts(count, frequency = freq, start = start_list)
  auto_var <- auto.arima(type_ts)
  type_ts_forecast1 <- forecast(auto_var, level = c(50, 68.2, 90, 95), h = nforecasts)
  
  # optionally produce forecast plots
  if(make_plot){
    plot(type_ts_forecast1, xlab = "year", ylab = type_val)
  }
  
  # turn forecasts into dataframe
  forecasts <- type_ts_forecast1 %>% 
    as.data.frame() %>% 
    rownames_to_column("date") %>% 
    mutate(type = type_val)

  # Rename forecasts columns
  colnames(forecasts) <- c("date", "count", "q_25", "q_75", "q_16", "q_84", "q_05", "q_95", "q_2p5", "q_97p5", "type")

  return(forecasts)
}
```

# Load function

```{r load dataframe and prep for forecast}
load_forecast_df <- function(city = "nashville"){
  #' Load dataframe for npermits forecast
  #'
  #' Loads the dataframe containing the historical permits we wish to forecast into the future. A few date manipulations are done to make this work with `forecast_npermits_sub`.
  #'
  #' @param city (char, default="nashville") City upon which forecast will be built. Only valid option for now is "nashville"
  #'
  #' @return modified dataframe that can be fed into `forecast_npermits_sub`
  
  if(city!="nashville"){
    stop("Only valid city for now is nashville")
  }
  
  load_permits(city="nashville") %>%
    filter(generates_debris) %>%        # only permits deemed to generate debris
    filter(comm_v_res!="other") %>%     # I suspect these should be removed
    mutate(fy = as.numeric(substr(date_issued, 1, 4)) + (substr(date_issued, 6, 7) >= "07")) %>%
    mutate(yrmonth = format(date_issued, "%Y%m")) %>%
    filter(yrmonth != "201606" & yrmonth != "202106")
}
```

# Test

```{r test funcs, purl=FALSE}
forecast_cols <- c("comm_v_res", "project_type")

#set to monthly
load_forecast_df() %>% 
  forecast_npermits_sub(forecast_cols, freq=12, make_plot=TRUE) %>% 
  arrange(year,month)

#set to annual
load_forecast_df() %>% 
  forecast_npermits_sub(forecast_cols, df_start_yr = 2017, df_start_mo = NA, freq = 1, make_plot=TRUE) %>% 
  arrange(year)
```

################################################################################
Everything below consists of original tests that have been superseded by the above functions. I'm leaving it for now, but it shouldn't be run. The chunks should contain `purl=FALSE` to avoid running when sourcing this file.

```{r nash_time_eda, purl=FALSE}

#takes nash permits and preps for time series

#loads older nash_permits that has yet to be cleaned but does include date_entered
nash_permits <- read_feather(expand_boxpath("Nashville/nash_permits.feather"))

nash_permits_count = nash_permits %>%
  filter(permit_subtype_description != "Sign - Ground /  Wall Signs") %>%
  filter(permit_type_description != "Building Commercial - Change Contractor" & permit_type_description != "Building Residential - Change Contractor") %>%
  mutate(project_type = if_else(permit_type == "CADM", "demolition", "construction")) %>%
  mutate(comm_v_res = if_else(str_detect(permit_type_description, "Residential"), "residential", "commercial")) %>%
  mutate(comm_v_res = if_else(str_detect(permit_subtype_description, "Resid"), "residential", comm_v_res)) %>%
  #should i filter out the cancelled stuff for forecast? doing so for now
  filter(status %in% c("DONE","EXPIRED","ISSUED","OK TO PAY","OPEN","PENDING")) %>%
  #assuming fiscal year starts July 1, below assigns Fy
  mutate(fy = as.numeric(substr(date_entered, 1, 4)) + (substr(date_entered, 6, 7) >= "07")) %>%
  #month
  mutate(yrmonth = format(date_entered, "%Y%m")) %>%
  mutate(type = paste(comm_v_res, project_type, sep=" "))
  
nash_permits_count_fy = nash_permits_count %>%
  group_by(fy, type) %>%
  mutate(count = n()) %>%
  select(fy, type, count) %>%
  distinct() %>%
  ungroup() %>%
  #only have 2 weeks of data from fy 2016, removing
  filter(fy != 2016)
  #as a note, missing 2 weeks of data from fy 2020

plot_fy = ggplot(data=nash_permits_count_fy, aes(x=fy, y=count, group=type)) +
  geom_line(aes(color=type))+
  geom_point(aes(color=type)) +
  ggtitle("Permits entered per fiscal year by type")

plot_fy

nash_permits_count_fy

######## now by month

nash_permits_count_mo = nash_permits_count %>%
  group_by(yrmonth, type) %>%
  mutate(count = n()) %>%
  select(yrmonth, type, count) %>%
  distinct() %>%
  ungroup() %>%
  #removing below 2 months because of missing 2 weeks of data in each month
  filter(yrmonth != "201606" & yrmonth != "202106")

plot_months = ggplot(data=nash_permits_count_mo, aes(x=yrmonth, y=count, group=type)) +
  geom_line(aes(color=type))+
  geom_point(aes(color=type)) +
  ggtitle("Permits entered per month by type") + 
  theme(axis.text.x = element_text(angle=90, hjust=1, size= 5))

plot_months

######## now by month, minus 6 months for testing

nash_permits_count_mo6 = nash_permits_count_mo %>%
  filter(yrmonth != "202105" & yrmonth != "202104" & yrmonth != "202103" & yrmonth != "202102" & yrmonth != "202101" & yrmonth != "202012")

plot_months6 = ggplot(data=nash_permits_count_mo6, aes(x=yrmonth, y=count, group=type)) +
  geom_line(aes(color=type))+
  geom_point(aes(color=type)) +
  ggtitle("Permits entered per month by type (minus 6 months)") + 
  theme(axis.text.x = element_text(angle=90, hjust=1, size= 5))

plot_months6

########


```


```{r forecast function original, purl=FALSE}

forecast_n_permits <- function(df, forecast_yrs = 5, df_start_yr = 2016, df_start_mo = 07, freq = 12) {
    
  months_to_forecast = forecast_yrs * freq
  start_list = c(df_start_yr, df_start_mo)
  start_list = start_list[!is.na(start_list)]
  
  comm_contr <- df %>% 
    filter(type== "commercial construction") %>%
    mutate(row = row_number())
  
  res_contr <- df %>% 
    filter(type== "residential construction") %>%
    mutate(row = row_number())
  
  comm_demo <- df %>% 
    filter(type== "commercial demolition") %>%
    mutate(row = row_number())
  
  res_demo <- df %>% 
    filter(type== "residential demolition") %>%
    mutate(row = row_number())

  forecast_by_subgroup <- function(type) {
  
    count = type$count
    count
    type_ts = ts(count, frequency = freq, start = start_list)
    #plot(type_ts)
    type_ts
    
    auto_var = auto.arima(type_ts)
    
    #type_ts_forecast1 = as.data.frame(forecast(auto_var))
    type_ts_forecast1 = forecast(auto_var, level = c(68.2, 95), h = months_to_forecast)
    return(type_ts_forecast1)

  }
  
  plot(forecast_by_subgroup(comm_contr),xlab = "commercial construction forecast")
  plot(forecast_by_subgroup(res_contr),xlab = "residential construction forecast")
  plot(forecast_by_subgroup(comm_demo),xlab = "commercial demolition forecast")
  plot(forecast_by_subgroup(res_demo),xlab = "residential demolition forecast")
  
  comm_contr_forecast = as.data.frame(forecast_by_subgroup(comm_contr)) %>%
    mutate(type="commercial construction") %>%
    rownames_to_column("date")
  
  res_contr_forecast = as.data.frame(forecast_by_subgroup(res_contr)) %>%
    mutate(type="residential construction") %>%
    rownames_to_column("date")
  
  comm_demo_forecast = as.data.frame(forecast_by_subgroup(comm_demo))%>%
    mutate(type="commercial demolition") %>%
    rownames_to_column("date")
  
  res_demo_forecast = as.data.frame(forecast_by_subgroup(res_demo))%>%
    mutate(type="residential demolition") %>%
    rownames_to_column("date")
  
  a = bind_rows(comm_contr_forecast, res_contr_forecast)
  b = bind_rows(comm_demo_forecast, res_demo_forecast)
  c = bind_rows(a,b)
  
  colnames(c) = c("date", "count","lo68","hi68","lo95","hi95", "type")
  
  d = c %>%
    mutate(sd = count-lo68) %>%
    select(date, type, count, sd) %>%
    mutate(data = "prediction")
  d
  
  df1 = df %>%
    mutate(sd = NA) %>%
    mutate(data = "observed")
  df1
  
  if(freq == 12){
    
    d = d  %>%
      mutate(date = as.Date(paste("01",date,sep="-"), format = "%d-%b %Y")) %>%
      mutate(date = format(as.Date(date), "%Y%m"))
    
    df1 = df1 %>% 
      #mutate(date = as.Date(paste("01",yrmonth,sep="-"), format = "%d-%Y%m")) %>%
      mutate(date = yrmonth) %>%
      select(date, type, count, sd, data)
    
    e = bind_rows(df1,d) %>%
      mutate(commercial = type == "commercial construction" | type == "commercial demolition") %>%
      mutate(construction = type == "residential construction" | type == "residential demolition")
    return(e)
  }
  
  if(freq == 1){
    df1 = df1 %>% 
      mutate(date = as.character(fy)) %>%
      select(date, type, count, sd, data)
    
    e = bind_rows(df1,d) %>%
      mutate(commercial = type == "commercial construction" | type == "commercial demolition") %>%
      mutate(construction = type == "residential construction" | type == "residential demolition")
    
    return(e)
  }
  
  return(d)

}


```



```{r original_fxn_tests, purl=FALSE}

forecast_mo <- forecast_n_permits(nash_permits_count_mo) 
#monthly forecast works with default options

forecast_n_permits(nash_permits_count_fy, df_start_yr = 2017, df_start_mo = NA, freq = 1) 
#options for annual forecasts (all years in fiscal years)

forecast_n_permits(nash_permits_count_mo6, forecast_yrs = 0.5) 
#options for annual forecasts (all years in fiscal years)

```



```{r check_v_6mo, purl = FALSE}

nash_permits_count_mo_select = nash_permits_count_mo %>%
  filter(yrmonth == "202105" | yrmonth == "202104" | yrmonth == "202103" | yrmonth == "202102" | yrmonth == "202101" | yrmonth == "202012") %>%
  mutate(data = "observed")
  
nash_permits_count_mo_select

predict6 = forecast_n_permits(nash_permits_count_mo6, forecast_yrs = 0.5)
predict6_form = predict6 %>%
  mutate(yrmonth = date) %>%
  filter(data == "prediction") %>%
  select(yrmonth,type,count, data)

predict6_form

nash_permits_count_forecast <- rbind(nash_permits_count_mo_select, predict6_form) %>%
  mutate(data_type = paste(type, data, sep=" "))
nash_permits_count_forecast

plot_all_f = ggplot(data=nash_permits_count_forecast, aes(x=yrmonth, y=count, group=data_type)) +
  geom_line(aes(color=data_type))+
  geom_point(aes(color=data_type)) +
  ggtitle("FY22 Forecast: Permits per FY by type")

plot_all_f


```


```{r rmse_calc, purl = FALSE}

rmse <- function (y_pred, y_true) 
{
  RMSE <- sqrt(mean((y_true - y_pred)^2))
  return(RMSE)
}
#result = rmse(actual, predicted)

nash_permits_count_rmse = left_join(nash_permits_count_mo_select, predict6_form, by=c("yrmonth","type")) %>%
  mutate(actual = as.numeric(count.x)) %>%
  mutate(predicted = as.numeric(count.y)) %>%
  select(-data.x,-data.y) %>%
  group_by(type) %>%
  summarise(
    RMSE = rmse(actual, predicted),
    R2 = cor(actual, predicted)^2
  )
  
nash_permits_count_rmse

#RMSE 23.9
```


```{r linear_reg, eval = FALSE, purl = FALSE}

####

#MODELING COMMERCIAL CONSTRUCTION

####

p_load(tidymodels)

comm_contr <- nash_permits_count_mo %>% 
  filter(type== "commercial construction") %>%
  mutate(row = row_number())

res_contr <- nash_permits_count_mo %>% 
  filter(type== "residential construction") %>%
  mutate(row = row_number())

comm_demo <- nash_permits_count_mo %>% 
  filter(type== "commercial demolition") %>%
  mutate(row = row_number())

res_demo <- nash_permits_count_mo %>% 
  filter(type== "residential demolition") %>%
  mutate(row = row_number())


comm_contr


count = comm_contr$count

#training testing set

set.seed(123)
splits      <- initial_split(comm_contr, strata = count)

comm_other <- training(splits)
comm_test  <- testing(splits)

#as val set

set.seed(234)
val_set <- validation_split(comm_other, 
                            strata = count, 
                            prop = 0.80)
val_set

#model

lr_mod <- 
  linear_reg(mode = "regression", penalty = tune(), mixture = 1) %>% 
  set_engine("lm")


#recipe

lr_recipe <- 
  recipe(count ~ row, data = comm_contr)

#

lr_wf <- workflow() %>%
  add_model(lr_mod) %>%
  add_recipe(lr_recipe) %>%
  fit(comm_contr)

lr_wf

pred= predict(lr_wf, comm_contr)

plot(pred$.pred)

pred

```
