---
title: "42-synthetic-permits"
output: html_notebook
---

```{r setup synthetic permits}
source(knitr::purl("30-feature-engineering.Rmd", output = "tmp-30.R", quiet=TRUE))
fs::file_delete("tmp-30.R")
p_load(docstring, glue, splitstackshape)
```

This notebook contains functions which generate features for synthetic permits. Each feature will be generated in a different way. At the moment, the only features I actually need to generate are `comm_v_res`, `project_type`, and `const_cost`. The former two are generated based on user-input values (though maybe will ultimately be predicted). To generate `const_cost`, I need to assume a distribution values and draw from it. Because the distributions of `const_cost` are different depending on `comm_v_res` and `project_type`, we want to use the appropriate distribution given those values.

Note -- docstrings for functions in this notebook can be viewed by running `docstring(function_name)`.

# Driver function

```{r generate synthetic permits}
gen_syn_permits <- function(num_permits_sub, df_source, extra_cols = NULL, count_col = "count", ind_num = FALSE, seed = NULL){
  #' Generate synthetic permits
  #'
  #' Generate a dataframe containing features of synthetic permits. The tibble `num_permits_sub` defines specific subtypes of interest and the number of desired permits for each subtype. The extra columns in `extra_cols` are generated via stratified bootstrap resampling of values in `df_source` (i.e., separately for each subtype defined in `num_permits_sub`).
  #'
  #' @param num_permits_sub (tibble) Categorical counts for each unique combination of subtypes
  #' @param df_source (tibble) Source data from which features will be sampled
  #' @param extra_cols (chr array, default = NULL) names of columns whose values will be generated via bootstrap resampling of `df_source`. If NULL all columns in `df_source` will be used.
  #' @param count_col (chr, default = "count") name of the column in `num_permits_sub` containing counts of each subtype. Assumed/default name is "count".
  #' @param ind_num (bool, default = FALSE) If TRUE, sampling of each extra column in `extra_cols` is done independently. If FALSE, entire rows of `df_source` are sampled, roughly preserving correlations. Default is FALSE
  #' @param seed (num, default = NULL) random seed
  #'
  #' @return (tibble) Synthetic permits
  #' @export
  #'
  #' @examples
  #' # After loading nash_permits
  #' 
  #' gen_syn_permits(num_permits_sub = tibble(project_type = c("construction", "construction", "demolition", "demolition"),
  #'                                          comm_v_res = c("commercial", "residential", "commercial", "residential"),
  #'                                          count = c(20, 5, 7, 2)),
  #'                 df_source = nash_permits)
  
  if(ind_num){
    stop(glue("The thing is...I haven't set up the code to independently bootstrap numerical columns."))
  }
  
  # set seed
  set.seed(seed)
  
  # get categorical columns
  forecast_cols <- num_permits_sub %>% 
    select(-!!count_col) %>% 
    colnames()
  
  # get columns to sample if not passed (from source dataframe)
  if(is.null(extra_cols)){
    extra_cols <- colnames(df_source)
    extra_cols <- extra_cols[!(extra_cols %in% forecast_cols)]
  }

  # Get counts for each subclass in a form usable by `stratified`
  combo_list <- num_permits_sub %>% 
    get_combo_list(count_col = count_col)

  # Generate synthetic permits by stratified resampling of source dataframe
  syn_permits <- df_source %>% 
    stratified(forecast_cols, combo_list, replace=TRUE) %>% 
    select(all_of(forecast_cols), all_of(extra_cols))
  
  return(syn_permits)
}
```


```{r get list of combined }
get_combo_list <- function(df, count_col = "count"){
  #' Get list of counts for categorical variable combinations
  #'
  #' Given a dataframe containing categorical variables and subtype counts (i.e., the output of `group_by(a, b)` followed by `summarise(count=n()`), return a named list where the names are the space-separated concatenations of the values in the grouped variables and the vaue is the counts. This list is useful for stratified resampling via `splitstackshape::stratified`.
  #'
  #' @param df (tibble) Categorical variables and counts for each grouped combo of variables.
  #' @param count_col (chr, default = "count") name of the column in `df` containing counts of each subtype.
  #'
  #' @return (list) Counts for categorical variable combinations (space-separated concatentations)
  #' @export
  #'
  #' @examples
  #' get_combo_list(df = tibble(project_type = c("construction", "construction", "demolition", "demolition"),
  #'                                             comm_v_res = c("commercial", "residential", "commercial", "residential"),
  #'                                             count = c(20, 5, 7, 2)))
  
  df %>% 
    unite(combo, !(!!count_col), sep=" ") %>% 
    pivot_wider(names_from = combo, values_from = {{count_col}}) %>% 
    as.list()
}
```


# Miscellaneous helper functions

## Load dataframes from which we construct distributions

```{r load dataframe for specific cities}
load_source_df <- function(city = "nashville"){
  #' Load permit dataframe for specified city
  #'
  #' Load in a dataframe with construction & demolition permit information, given the name of a valid city. If an invalid name is supplied no dataframe is loaded.
  #'
  #' @param city (chr, default="nashville') Name of city.
  #'
  #' @return (tibble) Features of construction/demoltion permits.
  #' @export 
  #'
  #' @examples
  #' load_source_df(city="nashville")
  
  if(city!="nashville"){
    stop("Temporarily, nashville is the only valid source city for synthetic permit generation")
  }
  # fname <- expand_boxpath("Nashville/adams_nashville_no_touching.feather")
  # df <- read_feather(fname)

  df <- load_permits(city = city) %>%
    filter(generates_debris) %>%      # only permits deemed to generate debris
    filter(comm_v_res!="other")       # I suspect these should be removed
    
  return(df)
}
```


## Test function
Here is a function to test whether or not the permit generation is working. Right now, I simply print out the number of permits generated of each subtype. This is pretty rudimentary. Eventually, adding some visualization or explicit checks would be nice.

```{r test syn permits}
run_sim_test <- function(){
  
  # load nashville
  nash_permits <- load_source_df()
  
  forecast_cols = c("comm_v_res", "project_type")
  extra_cols = c("const_cost","construction_subtype", "permit_subtype_description", "council_dist")
  
  num_permits_sub <- 
    nash_permits %>% 
    group_by(across(all_of(forecast_cols))) %>% 
    summarise(count = n()) %>% 
    ungroup()
  
  # generate synthetic permits
  syn_permits <- gen_syn_permits(num_permits_sub, nash_permits, extra_cols = extra_cols)
  
  syn_totals <- syn_permits %>% 
    group_by(across(all_of(forecast_cols))) %>% 
    summarise(count = n()) %>% 
    ungroup()
  
  # syn_totals and num_permits_sub should be the same
  if(identical(num_permits_sub, syn_totals)){
    print("Produced expected number of permits of each subtype. Looking good!")
  }
  else{
    stop("Something went wrong with the test. Did not produce expected number of permits of each subtype.")
  }
  
  return(syn_permits)
}
```


Call the test function.
```{r run test,purl=FALSE}
run_sim_test()
```

