---
title: "45-prepare-for-shiny"
output: html_notebook
---


# Prepare for Shiny

To reduce the amount of data involved in the Shiny app, we are performing all of the filters and computing summary statistics here, then saving the output.

Below, we add the desired cost categories. These need to be treated as a factor so that the `group_by` below can include empty groups when performing summary statistics (this is a peculiarity of how `group_by` works). If we do not have values for every possible filter combination, I believe the app will break.

```{r main prep function}
prepare_for_shiny <- function(permits_fname, forecast_fname){
  
  # Read data
  syn_permits <- read_feather(permits_fname) %>% 
    filter(fy>2021&fy<2027) %>% 
    mutate(cost_cat = as.factor(case_when(is.na(const_cost) ~ NA_character_,
                                          const_cost<50000 ~ "0-50k",
                                          const_cost<500000 ~ "50k-500k",
                                          const_cost<3000000 ~ "500k-3M",
                                          const_cost<20000000 ~ "3M-20M",
                                          TRUE ~ ">20M")))
  
  # Read/write forecast to correct location
  read_feather(forecast_fname) %>% 
    write_feather("nash-zero-shiny/shiny-data/best_forecast.feather")
  
  # debris summary
  group_cats <- c("comm_v_res","project_type","cost_cat")
  syn_permits %>% 
    summarise_shiny(group_cats, "fiscal_year") %>% 
    bind_rows(summarise_shiny(syn_permits, group_cats, "fiscal_quarter")) %>% 
    mutate(cost_cat = as.character(cost_cat)) %>% 
    mutate(across(all_of(group_cats), ~replace_na(.x, "All"))) %>% 
    mutate(comm_v_res = str_to_sentence(comm_v_res)) %>% 
    mutate(project_type = str_to_sentence(project_type)) %>% 
    filter(!(fiscal_year%in%c("FY 21", "FY 27"))) %>% 
    filter(!(fiscal_quarter%in%c("FY 21 - Q4", "FY 27 - Q1", "FY 27 - Q2", "FY 27 - Q3", "FY 27 - Q4"))) %>% 
    write_feather("nash-zero-shiny/shiny-data/syn_permit_summary.feather")
  
  # map summary
  group_cats <- c("comm_v_res","project_type")
  
  syn_permits %>% 
    summarise_shiny_map(group_cats, "fiscal_year") %>% 
    mutate(across(all_of(group_cats), ~replace_na(.x, "All"))) %>% 
    mutate(comm_v_res = str_to_sentence(comm_v_res)) %>% 
    mutate(project_type = str_to_sentence(project_type)) %>% 
    filter(fy>2021 & fy<2027) %>%
    write_feather("nash-zero-shiny/shiny-data/map_summary.feather")
}
```



This function below could probably be generalized, but I don't know if I want to. For now, it takes in the permits dataframe, a list of grouping categories, and a time quantity (fiscal_year or fiscal_quarter). It will return a summary (median and desired quantiles) for the requested groupings. (Summary stats are computed across samples.)

```{r debris summary function}
debris_summ <- function(df, group_cats, time){
  
  ts <- rlang::sym(time)
  df %>% 
    group_by(!!ts, sample, across(all_of(group_cats)),.drop = FALSE) %>% 
    summarise(debris = sum(total_debris, na.rm=TRUE)) %>% 
    ungroup() %>%
    group_by(!!ts, across(all_of(group_cats))) %>%
    summarise(total_debris = median(debris),
              q_05 = quantile(debris, 0.05),
              q_25 = quantile(debris, 0.25),
              q_75 = quantile(debris, 0.75),
              q_95 = quantile(debris, 0.95)) %>%
    ungroup()
}
```


```{r summary for shiny}
summarise_shiny <- function(df, group_cats, time){

  # all cats
  df_summ <- df %>% 
    debris_summ(group_cats, time)
  
  # no cats plus bind
  df_summ <- df_summ %>% 
    bind_rows(debris_summ(df, c(), time))
  
  # single and double groupings plus bind
  for(cat in group_cats){
    gcats <- group_cats[!(group_cats %in% c(cat))]
    df_summ <- df_summ %>% 
      bind_rows(debris_summ(df, cat, time)) %>% 
      bind_rows(debris_summ(df, gcats, time))
  }
  return(df_summ)
}
```

```{r map summary func}
map_summ <- function(df, group_cats, time){
  
  df %>% 
    group_by(fy, sample, council_dist, across(all_of(group_cats)), .drop = FALSE) %>%
    mutate(npermits = n()) %>% 
    mutate(across(c(total_debris, const_cost), mean, na.rm = TRUE, .names = "{.col}_mean")) %>% 
    mutate(total_debris = sum(total_debris, na.rm=TRUE)) %>% 
    ungroup() %>% 
    group_by(fy, council_dist, across(all_of(group_cats))) %>% 
    summarise(across(c(total_debris, total_debris_mean, const_cost_mean, npermits), mean, na.rm = TRUE)) %>% 
    ungroup()
}
```

```{r map summary shiny}
summarise_shiny_map <- function(df, group_cats, time){

  # all cats
  df_summ <- df %>% 
    map_summ(group_cats, time)
  
  # no cats plus bind
  df_summ <- df_summ %>% 
    bind_rows(map_summ(df, c(), time))
  
  # single and double groupings plus bind
  for(cat in group_cats){
    df_summ <- df_summ %>% 
      bind_rows(map_summ(df, cat, time))
  }
  return(df_summ)
}
```

